---
title: "regex exercise"
author: "DerekRBratcher"
date: "10/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://towardsdatascience.com/a-gentle-introduction-to-regular-expressions-with-r-df5e897ca432

stringr package functions
str_extract_all(string, pattern) this returns a vector of all pattern matches in a string
str_replace_all(string, pattern, replacement) this function returns the string with replacements where the pattern described was determined
```{r}
https://towardsdatascience.com/a-gentle-introduction-to-regular-expressions-with-r-df5e897ca432
basicString <- "Drew has 3 watermelons, Alex has 4 hamburgers, Karina has 12 tamales, and Anna has 6 soft pretzels"

captureGroup1 <- str_match_all(basicString, "[A-Z][a-z]+\\s\\w+\\s\\d{1,2}\\s\\w+\\s*\\w*")
captureGroup1
```

```{r}
 basicString <- "Drew has 3 watermelons, Alex has 4 hamburgers, Karina has 12 tamales, and Anna has 6 soft pretzels"
```
If I want to pull every instance of one person’s name from this string, I would simply pass the name and basic_string to str_extract_all()

```{r}
Drewsinbasicstr<-str_extract_all(basicString, "Drew")
```
returns a list with one element [[1]]

replace alex /w shawn
```{r}
replalex4shawn<-str_replace_all(basicString, "Alex", 'Shawn')
```

There are several stackable concepts that drive regex:
Character sets
Meta characters
Quantifiers
Capture Groups


Character Sets 
`"[aeiou]"` for vowels
`"[0-9]"`  match any digit
`"[a-z]"`  match any letter
`[a-zA-Z0-9]` match all of the above
`[^aeiou]` match anything but lowercase letters
`[^0-9]`   match anything but a digit
Character sets represent options inside of brackets, with regex matching only one of the options. 
```{r}
#extract all instances of vowels from basic string
str_extract_all(basicString, "[aeiou]")
```
Match a range of characters: c-j

```{r}
#extract all instances of c to j lowercase from basic string
str_extract_all(basicString, "[c-j]")
```
Character sets can contain everything from this section simultaneously, so something like [A-Ct-z7-9] is still valid. It will match every character from capital “A” to capital “C,” lowercase “t” to lowercase “z,” and 7 through 9.

##Meta Characters
Meta characters represent a type of character. They will typically begin with a backslash \. Since the backslash \ is a special character in R, it needs to be escaped each time it is used with another backslash. In other words, R requires 2 backslashes when using meta characters. Each meta character will match to a single character.

\\s: This meta character represents spaces. This will match to each space, tab, and newline character. You may also specify \\t and \\n for tab and newline characters respectively. 

Side note: this string does not have any tabs, but be cautious when looking for them. Many integrated development environments, IDE’s, have a setting that will replace all tabs with spaces while you are typing. 
```{r}
#extract all instances of a space in basic string
str_extract_all(basicString, "\\s")
```


\\w: This meta character represents alphanumeric characters. This includes all the letters a-z, capital and lowercase, and the numbers 0–9. This would be the equivalent of the bracket group [A-Za-z0-9], just much quicker to write. Take caution in remembering that the \\w meta character on its own only captures a single character, not entire words or numbers. 

```{r}
stringr::str_extract_all(basicString, "\\w")
```
\\d: This meta character represents numeric digits. when applied to picnic example, it only finds the digits in the string. You’ll notice that like bracket groups `"[0-9]"`, it picks up 5 numbers instead of the 4 we expect. This is because it is looking for each **individual digit**, not groups of digits. 


```{r}
str_extract_all(basicString, "\\d")
str_match_all(basicString, "\\d")
```

Quantifiers
As we saw in the previous section, a single meta character can have somewhat limited functionality. When it comes to words or numbers, we usually want to find more than 1 character at a time. This is where quantifiers come in.
quantifiers allow you to specify how many of a character you're expecting. they are always following the character that youve specified with one of the commands above
+ quantifies 1 or more matches.
```{r}
#This is a good point to mention that **regex looks for non-overlapping matches**. In this case, it looks at each B and the character that follows it. Since we used the `+` quantifier, it continues to match until it reaches the end of a group of B’s.
quantExample <- "B.BB.BBB.BBBB"
#extract "B" one or more matches
str_extract_all(quantExample, "B+")
#match "B" one or more matches in a list/matrix
str_match_all(quantExample, "B+")
# extract all characters, length 3 
str_extract_all(basicString, "[A-Za-z]{3}")
#corrected extract all letters specify length 4
str_extract_all(basicString, "[A-Za-z]{4}")
# extract all characters, length 1 or 2
str_extract_all(basicString, "\\d{1,2}")
# extract all digits in matrix, no specific length
str_match_all(basicString, "\\d+")
#use match to match vector values exactly, each of this is a set of characters separated by an equal sign
stri_match_all_regex('breakfast=eggs, lunch=pizza, dessert=icecream',
   '(\\w+)=(\\w+)')
```


`{}` quantifies a specific number or range of matches. When written like {2} it will match exactly 2 of the preceding character. We’ll see some interesting results.

```{r}
#extract exactly 2 instances with 2 of the value "B" 
quantExact2 = str_extract_all(quantExample, "B{2}")
quantExact2
#reading from right to left there are 4 instances of a pair of B's
```
we can use a character set with all upper and lower case letters, adding a + quantifier to it. This will find any length of alpha characters grouped together. Said another way, it finds all of the words. 

```{r}

str_extract_all(basicString, "[A-Za-z]+")
```
To find the quantity of each food item, we can use the \\d meta character and the quantifier {1,2}. This will find the groups of digits that are 1 or 2 characters long. 
```{r}
str_extract_all(basicString, "\\d{1,2}")
```
To find the quantity and name of each food item, we can combine quantifiers with meta characters. We know that each number has a food item directly after it, so we can just add on to the previous example. We know there is a space and a word (\\s\\w+) that could be followed by another word like how “soft pretzel” appears. To specify the second word might not be there, we can use the * quantifier with the second word. Just like that we have a list containing the quantity and name of every good at our picnic.
```{r}
#find quantity of each food in "basic string" combine meta characters 
str_extract_all(basicString, "\\d{1,2}\\s\\w+")
#extract all matches from basic string (digits, length 1 and 2, space, alphanumerics unspecified length)
str_extract_all(basicString, "\\d{1,2}\\s\\w+\\s*\\w*")
# extract all the following matches from basic string: digits, length 1 and 2 units, one space, letters alphanumeric one or more, spacers 0 or more and letters 0 or more

# * quantifies zero or more matches. This can be helpful when we are looking for something that may or may not be in our string
# + quantifies 1 or more matches. 
#regex reads your data and identifies each match string by string or byte by byte based on the pattern described
```
Capture Groups
Capture groups allow you to look for entire phrases and only return parts of them. 
in "basicstring" match each person’s name, what they are bringing, and how much of it they are bringing.
str_extract_all captures individual elements
str_match_all captures groups in a matrix
```{r}
#str_match_all in basic string : capital leters,lowercase, space, alphanumeric 1 or more, space, digit 1 or 2, space, aplhanumeric one or more, space 0 or more, aphanumeric 0 or more
str_match_all(basicString, "[A-Z][a-z]+\\s\\w+\\s\\d{1,2}\\s\\w+\\s*\\w*")
str_extract_all(basicString, "[A-Z][a-z]+\\s\\w+\\s\\d{1,2}\\s\\w+\\s*\\w*")
```

```{r}
captureGroup1 <- str_match_all(basicString, "[A-Z][a-z]+\\s\\w+\\s\\d{1,2}\\s\\w+\\s*\\w*")
captureGroup1

```

The regex we used in captureGroup1 is looking for a name, which starts with a capital letter and has any amount of lowercase letters after it ([A-Z][a-z]+). Then after a space it matches the pattern space, word, space \\s\\w+\\s. Next we are looking for a 1 to 2 digit number followed by a space and a word (\\d{1,2}\\s\\w+). You can see in the output each row of the matrix is a character string with the details for each person.

By using capture groups, we can return a matrix where each column contains a specific piece of information. We’ll create capture groups containing each name, quantity, and item. Capture groups are simply sections of the regex that you wrap in **parenthesis**.

```{r}
captureGroup2 <- str_match_all(basicString, "([A-Z][a-z]+)\\s\\w+\\s(\\d{1,2})\\s(\\w+\\s*\\w*)")
captureGroup2
```
The first column in the matrix has the entire regex, ignoring the capture groups. The remaining columns of the matrix each correspond to the capture groups we defined for name, quantity, and item.

