---
title: "regex exercise"
author: "DerekRBratcher"
date: "10/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

stringr package functions
str_extract_all(string, pattern) this returns a vector of all pattern matches in a string
str_replace_all(string, pattern, replacement) this function returns the string with replacements where the pattern described was determined

```{r}
 basicString <- "Drew has 3 watermelons, Alex has 4 hamburgers, Karina has 12 tamales, and Anna has 6 soft pretzels"
```
If I want to pull every instance of one person’s name from this string, I would simply pass the name and basic_string to str_extract_all()

```{r}
Drewsinbasicstr<-str_extract_all(basicString, "Drew")
```
returns a list with one element [[1]]

replace alex /w shawn
```{r}
replalex4shawn<-str_replace_all(basicString, "Alex", 'Shawn')
```

There are several stackable concepts that drive regex:
Character sets
Meta characters
Quantifiers
Capture Groups


Character Sets 
`"[aeiou]"` for vowels
`"[0-9]"`  match any digit
`"[a-z]"`  match any letter
`[a-zA-Z0-9]` match all of the above
`[^aeiou]` match anything but lowercase letters
`[^0-9]`   match anything but a digit
Character sets represent options inside of brackets, with regex matching only one of the options. 
```{r}
#extract all instances of vowels from basic string
str_extract_all(basicString, "[aeiou]")
```
Match a range of characters: c-j

```{r}
#extract all instances of c to j lowercase from basic string
str_extract_all(basicString, "[c-j]")
```
Character sets can contain everything from this section simultaneously, so something like [A-Ct-z7-9] is still valid. It will match every character from capital “A” to capital “C,” lowercase “t” to lowercase “z,” and 7 through 9.

##Meta Characters
Meta characters represent a type of character. They will typically begin with a backslash \. Since the backslash \ is a special character in R, it needs to be escaped each time it is used with another backslash. In other words, R requires 2 backslashes when using meta characters. Each meta character will match to a single character.

\\s: This meta character represents spaces. This will match to each space, tab, and newline character. You may also specify \\t and \\n for tab and newline characters respectively. 

Side note: our example string does not have any tabs, but be cautious when looking for them. Many integrated development environments, or IDE’s, have a setting that will replace all tabs with spaces while you are typing. 
```{r}
#extract all instances of a space in basic string
str_extract_all(basicString, "\\s")
```


\\w: This meta character represents alphanumeric characters. This includes all the letters a-z, capital and lowercase, and the numbers 0–9. This would be the equivalent of the bracket group [A-Za-z0-9], just much quicker to write. Take caution in remembering that the \\w meta character on its own only captures a single character, not entire words or numbers. 

