---
title: "HW_03"
author: "DerekRBratcher"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Homework 3
1. (2 points) Suppose we have a dataset A (see code below) where each column represents multiple measures of nitrogen concentration in a particular lake. We want to get the average value for each lake. Do this in two ways: a `for` loop and a vectorized function `colMeans()`.
```{r}
set.seed(12) # to be reproducible
A = matrix(data = 1:500, nrow = 50, ncol = 10)
colnames(A) = paste("lake", 1:10, sep = "_")
```
(sidenote) set seed- When you generate “random numbers” in R, you are actually generating pseudorandom numbers. These numbers are generated with an algorithm that requires a seed(integer argument) to initialize. Being pseudorandom instead of pure random means that, if you know the seed and the generator, you can predict (and reproduce) the output.
The purpose of the R `set.seed` function is to allow you to set a seed and a generator (with the kind argument) in R.
`set.seed` is to arrange to create a vector of pseudorandomly ordered numbers by the random number generator (RNG) assigned to the integer argument
you have to recall that seed "12" with set seed to recall that sequence of 50 numbers
```{r}
set.seed(12)
x<- rnorm(50)
```
lines 15-16 set the matrix parameters. elements 1:500, number of rows 50, number of columns 10, column names number 1-10 and lake separated by an _

one can use set.seed to pull random observations from the newly made dataset
```{r}
#set seed to assign the RNG to 45 when generating 5 random observations from dataset A 

set.seed(45)
rndm<- sample(1:nrow(A), 5)
rndm
# you can use the following to call the whole row of the random values
A[rndm,]
```

