---
title: "HW_03"
author: "DerekRBratcher"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Homework 3
1. (2 points) Suppose we have a dataset A (see code below) where each column represents multiple measures of nitrogen concentration in a particular lake. We want to get the average value for each lake. Do this in two ways: a `for` loop and a vectorized function `colMeans()`.
```{r}
set.seed(12) # to be reproducible
A = matrix(data = 1:500, nrow = 50, ncol = 10)
colnames(A) = paste("lake", 1:10, sep = "_")
```
(sidenote) set seed- When you generate “random numbers” in R, you are actually generating pseudorandom numbers. These numbers are generated with an algorithm that requires a seed(integer argument) to initialize. Being pseudorandom instead of pure random means that, if you know the seed and the generator, you can predict (and reproduce) the output.
The purpose of the R `set.seed` function is to allow you to set a seed and a generator (with the kind argument) in R.
`set.seed` is to arrange to create a vector of pseudorandomly ordered numbers by the random number generator (RNG) assigned to the integer argument
you have to recall that seed "12" with set seed to recall that sequence of 50 numbers
```{r}
set.seed(12)
x<- rnorm(50)
```
lines 15-16 set the matrix parameters. elements 1:500, number of rows 50, number of columns 10, column names number 1-10 and lake separated by an _

one can use set.seed to pull pseudorandom observations for the newly made dataset
```{r}
#set seed to assign the RNG to 45 when generating 5 random observations from dataset A 

set.seed(45)
rndm<- sample(1:nrow(A), 5)
rndm
# you can use the following to call the whole row of the random values
A[rndm,]
#for each column in matrix A print the avg of that column 
for(i in 1:ncol(A)){
  print(mean(A[ ,i]))
}

```
take matrix a nitrate conc (rows) vs lakes (cols) get avg for each lake in a row using `for` loop and vectorized funct `colmeans()`

```{r}
# syntax
for (variable in sequence){
  statements for each variable
}

for (i in x) {
  print(paste(i, "^2 = ", i ^ 2, sep = ""))
}
```

print and Cat
print returns various data forms to the console. it will read all data except character strings
cat transcribes code specific info involving characters

```{r}
my_string <- "This is \nan example string"    # Create example string
print(my_string)
cat(my_string)
my_string
```
print for data
cat for characters
notice print is invisibly called when I call the value "my_string"
print can be modified by "quote" (TRUE | FALSE depending on if you want output to be printed /w quotation marks)
"max.levels" if data involves factors, this allows one to determine how many levels to return
"digits" determines the number of sig figs to return
"na.print" to incorporate NA values
"zero.print" to determine how 0 is handled

Paste & Paste0
paste converts arguments to character strings and concatenates/joins them (use sep to add spaces) paste0 means no sep.
The collapse argument can define a value to use when joining those values to create a single string - or& or_

here we paste the class of elements K,D,L with 1-9 separated by | collapsed with & 
so the output will print one of the 3 letters separated by a | and joined as one string by &
```{r}
paste(c("k", "D", "L"), 1:9, sep = "|", collapse = "&")
```

2. (2 points) From the for loop lecture, we see the following example of using apply():
 
```{r}
x = array(1:27, dim = c(3, 3, 3))
print(x)
apply(X = x, MARGIN = c(1, 2), 
      FUN = paste, collapse = ", ")
```
translation-
make an array use elements 1-27, the dimensions are x-3,y-3,z-3
print x
apply to the array x specifying all rows 1, and all cols 2 (you can also subset for specific rows /w ( x[c(1, ), ] ) to indicate just row 1
apply function "paste" to determine spacing in this case separate /w ", "
(Margin=you can use 1 to apply to all rows and 2 to apply to all cols to specify a element by )
FUN= funtion applied (+,paste,sum,summary,min,max,range)

now use a `for` loop to do the same
```{r} 
#tuitorial loop on dfb
dfb <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
median(df$a)

output<- vector("double",length = ncol(dfb))
for (i in seq_along(dfb)) {
  output[[i]]<- median(dfb[[i]])
}
print(output)


#label each dimension of array 
x1 <- x[ 1:3, 1:3, 1]
x2 <- x[1:3, 1:3, 2]
x3 <- x[1:3, 1:3, 3]
i <- x[ 3, 3, 1:3]
for (i in x1) {
  for (i in x2) 
    for (i in x3) {
  print(paste(x1, x2, x3, sep = ", "))
    }
  }
  

x1 <- x[1:3, 1:3, 1]
x2 <- x[1:3, 1:3, 2]
x3 <- x[1:3, 1:3, 3]
for (i in x1) {
   for (i in x2) 
     for (i in x3) {
  print(paste(x1, x2, x3, sep = ", "))
    }
  }


```

```{r}
#designate i-k by dimension each by row then by col
result<-array(1:27,  dim = c(3,3,3))

for (i in x[1:3, 1:3, 1]) {
  result[i]= as.matrix( print(paste(i), dim = 3,3,3))
  for (j in x[1:3, 1:3, 2]){
    result[j]= as.matrix( print(paste(j) ))
    for (k in x[1:3, 1:3, 3]) {
      result[k]= as.matrix( print(paste(k) ))
      result<- paste(i, j, k, sep = ", ")
    }
  }}
  print(as.matrix( result))
                     

#why didnt this work?
```

2 points) The Fibonacci Sequence is the series of numbers that the next number is the sum of the previous two numbers: 0, 1, 1, 2, 3, 5, 8 … Use a for loop to get the first 30 numbers of the Fibonacci Sequence. This question should demonstrate the need for loops because there is no easy way to use vectorized functions in this case.

```{r}
#set up a numeric vector of 40 elements and assign to fibonacci
Fibonacci <- numeric(40)
#make the first two elements 1
Fibonacci[1] <- Fibonacci[2]<- 1
#set up loop applying formula at 3-30
for (i in 3:40)  {
  Fibonacci[i]<- Fibonacci[i - 2] + Fibonacci[i - 1]
  print("The First 40 Fibonaci numbers")
  print(Fibonacci)
}

```
4. (2 points) In the example data below, extract those ranking numbers with _regular expression_. The results should have the number(s) and . if it follows after the numbers immediately (i.e., `1., 12., 105., 105.3`, etc.). Remove empty strings from the final results. You should get 107 strings for your results.
add 54. Electronic - Tighten Up and 55. Enigma - Sadeness (Part 1)
```{r}
top105
top_105_ranking_numbers <- str_sub(top105,start = 1, end =  5)
top_105_ranking_numbers <- top_105_ranking_numbers |>  gsub(pattern = "[[:alpha:]]",replacement =  " ", x = top_105_ranking_numbers)
 top_105_ranking_numbers <- top_105_ranking_numbers |>  gsub(pattern = "[[:blank:]]",replacement =  "", x = top_105_ranking_numbers)
top_105_ranking_numbers<- str_replace( top_105_ranking_numbers,"@", "")
top_105_ranking_numbers<- str_replace( top_105_ranking_numbers,":", "")
top_105_ranking_numbers<- str_replace( top_105_ranking_numbers,"..", ".")
library(stringi)
top_105_ranking_numbers<- stri_remove_empty(top_105_ranking_numbers, na_empty = FALSE)
top_105_ranking_numbers<- sub("3..", "3.", top_105_ranking_numbers, fixed = TRUE)
top_105_ranking_numbers<- sub("5..", "5.", top_105_ranking_numbers, fixed = TRUE)

```

