---
title: "Hw4 unresolved"
author: "DerekRBratcher"
date: "10/22/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

2. (2 points) With the extracted data frame, create two new columns based on the Date column: month and day. month would be the month abbrevations from the Date column; day would be the numeric numbers from the Date column. Although you can use whatever approach to get this done (do not enter them by hand…), I suggest you try to practice regular expression here (sub() or stringr::str_extract()).


Why doesnt this work and what would work 
```{r}
library(dplyr)
Month = case_when( 
  starts_with(Date, "Se") ~ "September",
  starts_with(Date, "Oc") ~ "October",
  starts_with(Date, "No") ~ "November",
  starts_with(Date, "De") ~ "December"
  )
Schedule/materials_df <- cbind(SMdf[, c(1, 2)], Day, Month)
Schedule/materials_df<- 'SMdf'|> mutate(Day = str_sub(SMdf[1], start = -2, end = -1))
#create day columns
Schedule/materials_df<- 'S&Mdf' |>  mutate("S&Mdf$Month" = case_when( 
  starts_with(Date, "Se") ~ "September",
  starts_with(Date, "Oc") ~ "October",
  starts_with(Date, "No") ~ "November",
  starts_with(Date, "De") ~ "December"
  ))
Schedule/materials_df
#use regex to apply month an day entries to each column
#apply a seach for digits for date column then apply numbers to day column
#apply search for each month  then apply to month column
#use mutate or transmute case when script
#mutate only for nonstrings?
```
change col names
```{r}
# Change colname of one column
colnames(data)[colnames(data) == "Old_Name"] <- "New_Name"
 
# Change colnames of all columns
colnames(data) <- c("New_Name1", "New_Name2", "New_Name3")
 
# Change colnames of some columns
colnames(data)[colnames(data) %in% c("Old_Name1", "Old_Name2")] <- c("New_Name1", "New_Name2")

```

