---
title: "Lec_11_Loops"
author: "DerekRBratcher"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> #Loops

 <br>
###Loops repeat two steps: evaluation and execution. For evaluation, we tell the computer to test for a specific condition, and there are two outcomes of this evaluation: condition satisfied and not satisfied. Then for each scenario, we can let the computer know what to do respectively.###

R provides three types of loops
+ for
+ while
+ repeat

###Repeat

```{r}
# 1st set parameters
x<- 5
#then call for repeat of expression under condition
repeat 
  { if (x > 25) 
  break 
  else 
    print (x); x <- x + 5}
#check if x > 25, if so stop/break, otherwise repeat printing the outcome of the function x + 5 
```
Print first five multiples of 3 using repeat loop.
```{r}
D <- 3
repeat
{ if (D > 15)
  break
  else
    print(D); D <- D + 3}

```

###While

while loops repeat an expression "while" a condtion is true
```{r}
# values; 1st while, next condition, then expression
x<- 5
while(x <= 25) {
  print(x)
  x <- x+5
}

```
set values to be used x is 5
set condition while x is less than or equal to 25
set the then to run when cond is met, print x also increase x by 5 
this directs the loop to check if x is less than on equal to 25 then print then increase by 5 then repeat the loop with another comparisson to 25
```{r}
#while loop to calc sum of 1st 10 numbers
#store first number as a variable
uno = 1
# store sum of loop
total = 0
#set up while loop condition of operating
while(uno <= 10){
  #function to run on uno
  total = total + uno
  
  #increment uno by 1
  uno = uno + 1
}
print(total)
#loop should run 0+1,1+2,2+3... till uno reaches 10
```

first in setting out to calc sum of 1st 10 numbers I declared to variables, uno, my starting number and total, my sum of adding each number. my condition is within my while statement (uno has to be <= 10)
the while loop will start at uno adding to the total to get the new total increasing incrementally by 1 until uno is nolonger <= 10

you can also use `next` and `break` in the brackets. break determines when to stop the loop. next allows you to interrupt the process by skipping itterations or adding to condition (like picking from a vector for numbers over 5 but less than 9)
```{r}
x <- 5
# while loop to print odd number between 1 to 10
while (x <= 10) {
  #skip even numbers
  if (x %% 2 == 0) {
    x <- x + 1
    next
  }
  #print odd numbers
  print(x)
  #list initial condition
  x <- x + 1
}
```
while x is less than 10 (also add one if x is even) print x increased by increments of 1
This loop only prints the odd numbers in the range of 5 to 10. To do this, we have used an if statement inside the while loop to check if number is divisible by 2.
+ if number is divisible by 2, then its value is simply incremented by 1 and the iteration is skipped using the next statement.
+ if number is not divisible by 2, then the variable is printed and its value is incremented by 1.

