---
title: "HW_03"
author: "DerekRBratcher"
date: "10/7/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# __Homework 3__  


comments
For Q1, try to save the results as a vector next time. Print won't save results.
For Q2, I think you got it during office hour. But it is not the case on your pdf file... (-1 point)
For Q3, Don't use [[]] for vectors, it is for list.
For Q4, I don't think you need these many commands to do it...
For Q5, don't write the results in hard way, you can use top_105_ranking_numbers[duplicated(top_105_ranking_numbers)].


 1. (2 points) Suppose we have a dataset A (see code below) where each column represents multiple measures of nitrogen concentration in a particular lake. We want to get the average value for each lake. Do this in two ways: a `for` loop and a vectorized function `colMeans()`.
```{r}
set.seed(12) # to be reproducible
A = matrix(data = runif(n =1:500), nrow = 50, ncol = 10)
colnames(A) = paste("lake", 1:10, sep = "_")
```

   + find mean via `for` loop
```{r}
for(i in 1:ncol(A)){
  print(mean(A[ ,i]))
}
save results in a vector
```
 
 
  + find mean via `colMeans()`
 

```{r}
colMeans(A)
```











2. (2 points) From the for loop lecture, we see the following example of using apply():
 
```{r}
x = array(1:27, dim = c(3, 3, 3))
apply(X = x, MARGIN = c(1, 2), 
      FUN = paste, collapse = ", ")
```
 + Now, use `for` loops to get the same task done (hint: nested loops). The results should be the same. 

```{r}

#make empty matrix
y = matrix(0, nrow = 3, ncol = 3)
#make for loop where you fill the rows and the columns of the empty vector with your vectors corresponding to the rows and cols of array X
for(i in 1:nrow(x)){
  for (j in 1:ncol(x)){
    y[i,j]<- paste(x[i, j, ], collapse = ",")
  }}




result<-array(1:27,  dim = c(3,3,3))
x1 <- x[ 1:3, 1:3, 1:3]
x2 <- x[1:3, 1:3, 2]
x3 <- x[1:3, 1:3, 3]
for (i in x[1:3, 1:3, 1]) {
  for (j in x[1:3, 1:3, 2]) 
    for (k in x[1:3, 1:3, 3]) {
    }
  }
   result<- array(data =  print(as.array(paste(x1, x2, x3, sep = ", "))), dim = c(3,3,3))
result


#how do i print only once
   #why does i,j,k = 9L,18L, 27L?

```




3. (2 points) The Fibonacci Sequence is the series of numbers that the next number is the sum of the previous two numbers: 0, 1, 1, 2, 3, 5, 8 … Use a for loop to get the first 30 numbers of the Fibonacci Sequence. This question should demonstrate the need for loops because there is no easy way to use vectorized functions in this case.

```{r}
#set up a numeric vector of 40 elements and assign to fibonacci
Fibonacci<- numeric(30)
#make the first two elements 1
Fibonacci[[1]]<- Fibonacci[[2]]<- 1
#set up loop applying formula at 3-30
for (i in 3:length(Fibonacci)) { 
  Fibonacci[i]<- Fibonacci[i-2] + Fibonacci[i-1]
}
print("The First 30 Fibonaci numbers")
 print(Fibonacci)

```
4. (2 points) In the example data below, extract those ranking numbers with _regular expression_. The results should have the number(s) and . if it follows after the numbers immediately (i.e., `1., 12., 105., 105.3`, etc.). Remove empty strings from the final results. You should get 107 strings for your results. 
add 54. Electronic - Tighten Up and 55. Enigma - Sadeness (Part 1)
```{r}
top105 = readLines("http://www.textfiles.com/music/ktop100.txt")
top105 = top105[-c(64, 65)] # missing No. 54 and 55
top105
top_105_ranking_numbers <-stringr::str_sub(top105,start = 1, end =  5)
top_105_ranking_numbers <- top_105_ranking_numbers |>  gsub(pattern = "[[:alpha:]]",replacement =  " ", x = top_105_ranking_numbers)
top_105_ranking_numbers<- sub("@", "", top_105_ranking_numbers, fixed = TRUE)
 top_105_ranking_numbers <- top_105_ranking_numbers |>  gsub(pattern = "[[:blank:]]",replacement =  "", x = top_105_ranking_numbers)
top_105_ranking_numbers<- stringr::str_replace_all( top_105_ranking_numbers,":", "")

library(stringi)
top_105_ranking_numbers<- stringi::stri_remove_empty(top_105_ranking_numbers, na_empty = FALSE)
top_105_ranking_numbers<- sub("3..", "3.", top_105_ranking_numbers, fixed = TRUE)
top_105_ranking_numbers<- sub("5..", "5.", top_105_ranking_numbers, fixed = TRUE)
top_105_ranking_numbers
```


```{r} 
#alt filter strategy

#use stringr and extract to search for the digits by seaching for traits shared only by digits
alt<-stringr::str_extract((top105), pattern = "^\\d+\\.\\d?")
na.omit(alt)
alt[!is.na(alt)]

#trim off (.)
alt2<- gsub(pattern = "\\.$", replacement = "", x1)
as.numeric(alt2)
#return only duplicated values
alt2[duplicated((alt2))] 
#sort by duplicates
sort(table(alt2), decreasing = FALSE)
#id duplicates
names(alt3)[alt3 ==2]
```

5. (2 points) For the vector with length of 107 you got from question 4, remove all trailing .. (hint: `?sub`). Then convert it to a numeric vector and find out which numbers have duplications (i.e., a tie in ranking). Don’t count by eyes, use R to find it out (hint: `table(), sort(); or duplicated(), which(), [` subsetting; there are more than one way to do so).

```{r}
top_105_ranking_numbers<- as.numeric(top_105_ranking_numbers)
table(duplicated(top_105_ranking_numbers))
table(top_105_ranking_numbers)

```

 duplicates @ 83,91,97.