---
title: "HW_03"
author: "DerekRBratcher"
date: "10/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Homework 3
1. (2 points) Suppose we have a dataset A (see code below) where each column represents multiple measures of nitrogen concentration in a particular lake. We want to get the average value for each lake. Do this in two ways: a `for` loop and a vectorized function `colMeans()`.
```{r}
set.seed(12) # to be reproducible
A = matrix(data = 1:500, nrow = 50, ncol = 10)
colnames(A) = paste("lake", 1:10, sep = "_")
```

find mean via `for` loop
```{r}
for(i in 1:ncol(A)){
  print(mean(A[ ,i]))
}
```
 find mean via `colMeans()`
 

```{r}
colMeans(A)
```
2. (2 points) From the for loop lecture, we see the following example of using apply():
 
```{r}
x = array(1:27, dim = c(3, 3, 3))
apply(X = x, MARGIN = c(1, 2), 
      FUN = paste, collapse = ", ")
```
Now, use `for` loops to get the same task done (hint: nested loops). The results should be the same. 

```{r}
x1 <- x[ 1, 1, 1:3]
x2 <- x[2, 2, 2:3]
x3 <- x[3, 3, 3]
for (i in x1) {
  for (i in x2) 
    for (i in x3) {
  print(paste(x1, x2, x3, sep = ", "))
    }
  }
  

#how do i direct my loop to print by dimension??
```



(2 points) The Fibonacci Sequence is the series of numbers that the next number is the sum of the previous two numbers: 0, 1, 1, 2, 3, 5, 8 … Use a for loop to get the first 30 numbers of the Fibonacci Sequence. This question should demonstrate the need for loops because there is no easy way to use vectorized functions in this case.

```{r}
#set up a numeric vector of 40 elements and assign to fibonacci
Fibonacci<- numeric(40)
#make the first two elements 1
Fibonacci[1]<- Fibonacci[2]<- 1
i<- Fibonacci[1]
#set up loop applying formula at 3-30
for (i in seq_along(Fibonacci)) { 
  Fibonacci[i]<- Fibonacci[i-2] + Fibonacci[i-1]
  print("The First 30 Fibonaci numbers")
  print(Fibonacci)
}


```


 