---
title: "hw_02"
author: "DerekRBratcher"
date: "10/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, include = TRUE, eval = TRUE)
```

1. (2 points) Fix each of the following common data frame subsetting errors:

```{r}
library(tidyverse,dplyr)
#fix mtcars[mtcars$cyl < 6]
subset(mtcars, cyl < 6)
mtcars[mtcars$cyl < 6, ]
# mtcars[-1:3,]
mtcars[-c(1:3),]
# mtcars[mtcars$cyl = 8, ]
mtcars[mtcars$cyl == 8, ]
# mtcars[mtcars$cyl == 4 | 6, ]
mtcars[mtcars$cyl != 8, ]
mtcars[mtcars$cyl == c(4, 6), ]
```
2. (1 point) Why does the following code generated five missing values?
`x = 1:5; x[NA]`
NA in, NA out 
I created a vector using numbers 1-5 then applied the logical constant NA to vector X
the above commands executed the equivalent of `is.na<- x` setting all elements of x (from first argument) to NA(with second argument) 

this is subsetting  and recycling not assigning
is there a way to view code in r lang via github?

3. (2 points) Why does `mtcars[1:15]` return an error? How does it differ from `mtcars[1:15, ]`?
the first argument `mtcars[1:15]` specifies the dataset but not the variable/observation. the first argument lacks a comma designating 1-15 as the specific observations or variables depending on placement.  `mtcars[1:15, ]` specifies selection of the first 15 observations of the mtcars dataset along with their corresponding variables

4. (2 points) Explain how does the following code work.

```{r}
x = matrix(c(1:3, NA, 5:7, NA, NA), nrow = 3)
x[is.na(x)] = 0

```
In R a matrix must be a two dimensional data strucutre of homgenous elements. first x was set as a matrix with 3 rows and based on the elements provided the length was inferred by prioritizing the dimensions then the data. with 9 elements and three rows specified 3 columns was determined. because byrow wasnt specified (`byrow = TRUE`)the matrix was filled columnwise.
then with `is.na(x)` all NA values of matrix X were designated as zero

5. (3 points) Load the Car Road Tests dataset (in R, run `data("mtcars"), ?mtcars)`, then add a new column named as `mpg_2` for the mtcars data frame. You can use if ... else ... or ifelse or any other functions that can get the job done. This new column will categorize mpg into four categories using the thresholds below:mpg_2
(low- mpg<16, low_intermediate 16 <= mpg < 21, intermediate_high 21 <= mpg < 26,high 26 <= mpg)

```{r}
mtcars1<- mtcars |> dplyr::mutate (mpg_2 = dplyr::case_when(
  mtcars$mpg < 16 ~ "low",
   mtcars$mpg < 21  ~ "low_intermediate",
   mtcars$mpg < 25  ~ "intermediate_high",
    mtcars$mpg >= 26 ~ "high",
  TRUE ~ "A"
))
glimpse(mtcars1)
```

```{r}

mtcars3 <- mtcars |> dplyr::mutate( mpg2= (mtcars$mpg))
mtcars3$mpg2 <-
  ifelse(mtcars3$mpg2 < 16, "low",
ifelse(mtcars3$mpg2 < 20, "low_intermediate", 
ifelse(mtcars3$mpg2 < 25, "intermediate_high",
ifelse(mtcars3$mpg2 < 34, "high", mtcars3$mpg2))))
glimpse(mtcars3)
```




