---
title: "hw_02"
author: "DerekRBratcher"
date: "10/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. (2 points) Fix each of the following common data frame subsetting errors:

```{r}
#fix mtcars[mtcars$cyl < 6]
subset(mtcars, cyl < 6)
mtcars[mtcars$cyl < 6, ]
# mtcars[-1:3,]
mtcars[-c(1:3),]
# mtcars[mtcars$cyl = 8, ]
mtcars[mtcars$cyl == 8, ]
# mtcars[mtcars$cyl == 4 | 6, ]
mtcars[mtcars$cyl != 8, ]
mtcars[mtcars$cyl == c(4, 6), ]
```
2. (1 point) Why does the following code generated five missing values?
`x = 1:5; x[NA]`
NA in, NA out
I created a vector using numbers 1-5 then applied the logical constant NA to vector X
the above commands executed the equivalent of `is.na<- x` setting all elements of x (from first argument) to NA(with second argument)


3. (2 points) Why does `mtcars[1:15]` return an error? How does it differ from `mtcars[1:15, ]`?
the first argument `mtcars[1:15]` specifies the dataset but not the variable/observation. the first argument lacks a comma designating 1-15 as the specific observations or variables depending on placement.  `mtcars[1:15, ]` specifies selection of the first 15 observations of the mtcars dataset along with their corresponding variables

4. (2 points) Explain how does the following code work.

```{r}
x = matrix(c(1:3, NA, 5:7, NA, NA), nrow = 3)
x[is.na(x)] = 0

```
In R a matrix must be a two dimensional data strucutre of homgenous elements. first x was set as a matrix with 3 rows and based on the elements provided the length was inferred by prioritizing the dimensions then the data. with 9 elements and three rows specified 3 columns was determined. because byrow wasnt specified (`byrow = TRUE`)the matrix was filled columnwise.
then with `is.na(x)` all NA values of matrix X were designated as zero
